
---
title: "Lab 2 Group"
author: "Henry Siegler, Chau Nguyen, Joshua Havstad, Beatrice Weier, Amrit Pradhan"
date: '2022-04-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
library(here)
library(leaflet)
library(rgdal)
```

# Part One: Identifying Bad Visualizations

## Disseting a Bad Visualization

__1__

The graph is showing us the percentage of people who believe that vaccines are safe for many different countries across the world. The countries are grouped into different major regions of the world, and each of these regions has a line that indicates the country that is the region's median. I think the authors of the graph meant to demonstrate how people of different regions of the world compare in how they feel about vaccines in general. I think they also meant to convey how individual countries compare to other countries in their region for this statistic.

__2__

The variables that appear to be displayed in the visualization are the percent of people who believe vaccines are safe, country, and global region. 

__3__

The aesthetic for the x axis is the percent of people who believe that vaccines are safe, the y axis aesthetic is global region, and there is a label aesthetic for country name.

__4__

I would call this graph a modified box plot because the points are only distributed for the x axis, which is a numeric variable. I would say it is modified because instead of showing the IQR and minimum and maximum values, it only shows the median value of the region and each individual country's value.


__5__

The main problem with the graph is that the points seem to be graphed to a y axis value as well because the point increase in the y axis value as the percent of people who believe vaccines are safe increases. The y axis value does not represent anything, and it only makes the graph confusing to interpret. I would completely remove this increase in the y axis altogether. I would remove the legend at the bottom of the graph because the region name is displayed next to each region. Also, the graph only shows the country name for the countries that are close to the minimum or maximum value for the region, and they are sometimes very close together. I would remove these country labels because it makes the graph harder for the reader to analyze.

## Improving a Bad Visualization

```{r}
survey <- read_excel(here::here("wgm2018-dataset-crosstabs-all-countries.xlsx"), 
                     sheet = "Full dataset")
```

```{r}
vaccines_data <- survey %>% 
  filter(Q25 != 99) %>% 
  select(WP5, Regions_Report, Q25) %>% 
  mutate(safe = case_when(Q25 %in% 1:2 ~ 1,
                          TRUE ~ 0),
         region = case_when(Regions_Report %in% c(10, 11, 12, 18) ~ "Asia",
                            
                            Regions_Report %in% c(3, 13) ~ "Middle East and North Africa",
                            
                            Regions_Report %in% c(1, 2, 4, 5) ~ "Sub-Saharan Africa",
                            
                            Regions_Report %in% 6:8 ~ "Americas",
                            
                            Regions_Report %in% c(15, 16, 17 , 0) ~ "Europe",
                            
                            Regions_Report %in% c(9, 14) ~ "Former Soviet Union"
                            ))
vaccines_data <- vaccines_data %>% 
  group_by(WP5, region) %>% 
  summarise(pct = mean(safe) * 100) %>% 
  mutate(region = as.factor(region))
```

```{r}
vaccines_data %>% 
  ggplot(aes(x = pct, y = reorder(region, pct, median))) + 
  geom_boxplot() + 
  geom_jitter(aes(color = region)) +
  theme_classic() + 
  scale_color_discrete() + 
  scale_x_continuous(breaks = seq(30, 100, by = 10),
                     labels = function(x) paste0(x, "%")) +
  ggtitle("Percent of People who Believe Vaccines are Safe") + 
  labs(subtitle = "Points represent individual countries") +
  theme(legend.position = "none",
        axis.title = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank())
```

# Part Two: Broad Visualization Improvement

## Second Data Visualization Improvement


__1__ Select a data visualization in the report that you think could be improved. Be sure to cite both the page number and figure title. Do your best to tell the story of this graph in words. That is, what is this graph telling you? What do you think the authors meant to convey with it?

The choropleth on page 65 of the report can be improved. This graph uses varying shades of the color light green to represent the level of confidence that people in particular countries have in health clinics and hospitals in their countries. 

__2__ List the variables that appear to be displayed in this visualization.
The variables that appear to be displayed in the visualization are country and level of confidence in hospitals and health care clinics. 

__3__ Now that you're versed in the grammar of graphics (ggplot), list the aesthetics used and which variables are specified for each.
The color of the countries is an aesthetic. The x aesthetic is the percent. The world map is a geometry.

__4__ What type of graph would you call this?

This is a choropleth.

__5__ List all of the problems or things you would improve about this graph.  

Green only is not an accessible color for red green color blind people. This is particularly true for this graph because it is just uses slighly different shades of light green. I would change the colors of the choropleth to make it more accessible. I would also fix the label of the legend to make it more descriptive and accurate. 


__6__ Improve the visualization above by either re-creating it with the issues you identified fixed OR by creating a new visualization that you believe tells the same story better.

## Improving the Bad Visualization

```{r}
country_diction <- read_excel('wgm2018-dataset-crosstabs-all-countries.xlsx',
                               sheet = "Data dictionary")
```

```{r}
#create a df with the names of the countries in the whack order that it came in
dictionary_clean <- country_diction %>% 
  select(`Variable Type & Codes*`) %>% 
  head(n = 1) %>%  
  pull()
dictionary_clean <- dictionary_clean %>% 
  str_replace_all("[:digit:]+=", "") %>% 
  str_split(",") %>% 
  as.data.frame() %>% 
  mutate(across(where(is.character), str_trim)) %>%
  rename("countries" = "c..United.States.....Egypt.....Morocco.....Lebanon.....Saudi.Arabia...") %>% 
  head(n = 144) # last entry was blank so I removed it
```

```{r, warning = FALSE}
#clean data
trust_data <- survey %>% 
  group_by(WP5, Regions_Report) %>% 
  filter(Q22 != 98:99) %>% 
  select(WP5, Regions_Report, Q22) %>% 
  mutate(Q22 = case_when(Q22 %in% 1:2 ~ 1, 
                         TRUE ~ 0)) %>% 
  drop_na()
#calculate %
trust_data <- trust_data %>% 
  group_by(WP5) %>% 
  summarise(percent_trust = mean(Q22) * 100)
#add country column to data set
trust_data <- add_column(trust_data, dictionary_clean)
trust_data <- trust_data %>% 
  rename("NAME" = 'countries')
```

The leaflet base map came from this website: https://r-graph-gallery.com/183-choropleth-map-with-leaflet.html

Info on merging the data frame and the spacial object here:
https://gis.stackexchange.com/questions/259803/merge-spatial-and-non-spatial-data-and-create-spatialpolygonsdataframe-in-r

Some of the code for the leaflet map came from here:
https://slcladal.github.io/maps.html

```{r}
#read in spacial object
world <- readOGR(dsn = here::here("TM_WORLD_BORDERS_SIMPL-0.3", 
                                  "TM_WORLD_BORDERS_SIMPL-0.3.shp"))
#merge spacial object with df
worldmap <- merge(world, trust_data)
```

```{r}
# define colors
pal <- colorQuantile(rev(viridis::viridis(10)),
                      worldmap$percent_trust, n=10)
labels <- paste("<p>", worldmap$NAME, "<p>",
                "<p>", "%:", round(worldmap$percent_trust, digits = 2), "<p>",
                sep = "")
#leaflet map
leaflet(worldmap, 
        options = leafletOptions(attributionControl = FALSE)) %>%
  addTiles() %>% 
  setView(lat = 20, lng = 0, zoom = 1) %>%
  addPolygons(label = lapply(labels, htmltools::HTML), 
              labelOptions= labelOptions(direction = 'auto'), 
              weight=1, 
              color='white', 
              opacity=1,
              fillColor = ~pal(percent_trust), 
              fillOpacity = 1,
              highlightOptions = highlightOptions(
                color='#000000', 
                weight = 2,
                bringToFront = TRUE, 
                sendToBack = TRUE)) %>%
  addLegend("topright", 
            pal = pal, 
            values = ~percent_trust, 
            title = htmltools::HTML("Trust in Health Care Clinics"), 
            opacity = .8 )
```

## Third Data Visualization Improvement

__1__ Select a data visualization in the report that you think could be improved. Be sure to cite both the page number and figure title. Do your best to tell the story of this graph in words. That is, what is this graph telling you? What do you think the authors meant to convey with it?

Chart 4.12 on page 95 could be improved. This graph aims to visualize the perceived impact of technology on jobs based on the education level and socioeconomic status of countries. 

__2__ List the variables that appear to be displayed in this visualization.
The variables that appear to be displayed in this visualization are socioeconomic status of a country, perceived net impact that technology will have on jobs, and education status. 

__3__ Now that you're versed in the grammar of graphics (ggplot), list the aesthetics used and which variables are specified for each.
Education status is the fill aesthetic. Income level is the x aesthetic. The net impact score is the y aesthetic. The bar plot is the geometry. There is also a numeric label on top of all the bars (and under the upside down one); this is the label aesthetic.

_4__ What type of graph would you call this?
This is a bar plot that has one upside down bar.

__5__ List all of the problems or things you would improve about this graph.  
It is a confusing plot to understand. I have no idea how perceived net impact is measured just by looking at the plot. I have no idea what perceived net impact means either. It is a pretty confusing measure to try to understand just by viewing the plot. That should have been made clear. I feel like there should be more income groupings for countries as well because in the full dataset, countries are group in terms of being either high income, middle high income, middle low income, or low income. I think that a middle income grouping should be included. The colors are also not very nice looking.

## Improving the Bad Visualization
```{r, warning=FALSE}
#clean data
tech_data <- survey %>% 
  group_by(WP5, WBI, Education) %>% 
  select(WP5, Regions_Report, Q19) %>% 
  mutate(Q19 = case_when(Q19 == 1 ~ 1, 
                         Q19 %in%  2:3 ~ 0)) %>% 
  drop_na()
#calculate %
tech_data <- tech_data %>% 
  group_by(WP5, WBI, Education,) %>% 
  summarise(percent_job = mean(Q19) * 100, .groups = "keep")
#create groupings
tech_data <- tech_data %>% 
    mutate(
         income_level = case_when(WBI == 1 ~ "Low Income", 
                                  WBI %in% 2:3 ~ "Middle Income",
                                  WBI == 4 ~ "High Income"), 
         Education = case_when(Education == 1 ~ "Primary", 
                               Education == 2 ~ "Secondary", 
                              Education == 3 ~ "Post-secondary"))

#make income & education levels factors
tech_data <- tech_data %>%   
         mutate(income_level = factor(income_level, c("Low Income", "Middle Income", "High Income")), 
                Education = factor(Education, c("Primary", "Secondary", "Post-secondary")))
  
```

```{r}
#ridges plot
library(ggridges)
tech_data %>% 
  ggplot(mapping = aes(x = percent_job, y = income_level, fill = income_level)) + 
  geom_density_ridges() + 
  facet_wrap(~Education)+
  scale_fill_manual(values = c('#a8e6cf', '#dcedc1', '#ffd3b6', '#ffaaa5', '#ff8b94', '#bdeaee'))+
  scale_x_continuous(labels = function(x) paste0(x, "%")) + # Add percent sign 
  labs(x = "Percent that Believes Technology will Increase Local Jobs",
       subtitle = "Based on the Country's Income Level and Education Level") +
  theme_classic() +
  theme(legend.position = 'none',
        axis.title.y = element_blank()) +
  ggtitle("Percent of Respondents that \nBelieve Technology will Increase Jobs ")
```
